---
title: "Assignment 4"
author: "Tushar Ponkshe, tvp2110"
date: "October 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
setwd("C:/Users/tusha/OneDrive/Documents/Columbia Fall 2018 Courses/STAT COMP & INTRO TO DATA SCIENCE/Datasets")
debt <- read.csv("debt.csv", as.is = TRUE)
dim(debt)
head(debt)
```

**1.**
```{r}
#signif(debt, digits = 3)
#a
mean.growth = function(x) {
  mean(x[,"growth"])
}
signif(mean.growth(debt),3)


#b
require(plyr)
signif(daply(debt, .(Country), mean.growth), 3)

#Checking results
#mean(debt$growth[debt$Country=="Australia"])
```

**2.**
```{r}
library(ggplot2)
yearly.growth = daply(debt, .(Year), mean.growth)
signif(yearly.growth, 3)
Years = unique(debt$Year)
#ggplot(yearly.growth, aes(x = Years, y = mean.growth)) + geom_line()
plot(Years, yearly.growth, xlab = "year", ylab = "growth rates")
```

**3.**
```{r}
#a
#newdebt = na.omit(debt)
func = function(x) {
  cor(x$growth, x$ratio)
}

signif(func(debt), 3)
#b
cor.country = daply(debt, .(Country), func)
signif(head(cor.country), 3)
signif(mean(cor.country),3)

hist(cor.country, breaks = 10, xlab = "Correlation coefficient by country")

#c
cor.year = daply(debt, .(Year), func)
signif(head(cor.year), 3)
signif(mean(cor.year), 3)
hist(cor.year, xlab = "Correlation coefficient by year")
```

#d
The correlation coefficient of correlation for Norway seems to be very high compared to the rest  

**4.**
```{r}
fit = lm(debt$growth~debt$ratio)
signif(fit$coefficients, 3)

plot(debt$ratio, debt$growth, ylab = "GDP growth", xlab = "debt ratio")
abline(fit, col = 2)
```

**5.**
```{r}
four.countries = cor.country[which(cor.country < -0.5)]
signif(four.countries, 3)

by.country = split(debt, debt$Country)

par(mfrow=c(2,2))
plot(by.country$France$ratio, by.country$France$growth, main = "France", xlab = "debt ratio", ylab = "GDP growth")
plot(by.country$Germany$ratio, by.country$Germany$growth, main = "Germany", xlab = "debt ratio", ylab = "GDP growth")
plot(by.country$Italy$ratio, by.country$Italy$growth, main = "Italy", xlab = "debt ratio", ylab = "GDP growth")
plot(by.country$Japan$ratio, by.country$Japan$growth, main = "Japan", xlab = "debt ratio", ylab = "GDP growth")

```

**6.**
```{r}
library(dplyr)
#a
debt.france = by.country$France
dim(debt.france)

#b
debt.france$next.growth = mutate(debt.france, next.growth = ifelse((lead(Year) - Year) == 1, lead(growth), NA))
debt.france[debt.france$Year=="1971","next.growth"]
```

**7**
```{r}
func1 = function(df) {
  df$next.growth = mutate(df, next.growth = ifelse((lead(Year) - Year) == 1, lead(growth), NA))
}
debt.next = ddply(debt, .(Country), func1)
debt.next[(debt.next$Country=="France" & debt.next$Year=="2009"),]
```

**8**
```{r}
fit2 = lm(debt.next$next.growth~debt$ratio)
signif(fit2$coefficients,3)
summary(fit2)

plot(debt$ratio, debt.next$next.growth, ylab = "Next year's GDP growth", xlab = "This year's debt ratio")
abline(fit2, col=4)
```

How do they compare to the regression of the current year's growth on the current year's debt ratio?
- The coefficients for the intercept and the slope for the two linear models are very close in values.

**9**
```{r}
fit3 = lm(debt.next$next.growth~debt$growth)
signif(fit3$coefficients, 3)
plot(debt$growth, debt.next$next.growth, ylab = "Next year's GDP growth", xlab = "This year's GDP growth")
abline(fit3, col = 4)
summary(fit3)
```
Current Year's growth is a better predictor for future growth as seen from R squared in summary of the two fits